---
title: "Volatile Organic Compound (VOC) Analysis and Visualizations for Omniscent Data"
author: "Lourdes Vera"
output: word_document
editor_options: 
  chunk_output_type: inline
---
**The VOC data used in this script are from the Omniscent monitor and include benzene and toluene. They do not contain low confidence or false measurements.**

**The wind data, including windspeed and wind direction, were obtained using the Weather Underground webscraper, which can be found here: X** 

# Set Up
```{r setup, include=FALSE}
#Set up the R markdown file to allow it to "knit" into a document
knitr::opts_chunk$set(echo = TRUE)
```
## Load Libraries
```{r}
library(datetime)
library(lubridate)
library(ggplot2)
library(openair)
library(tidyverse)
library(showtext)
```
## Load Data
```{r}
benzene <- read.csv("Omniscent_Data - Benzene.csv", stringsAsFactors = FALSE)
toluene <- read.csv("Omniscent_Data - Toluene.csv", stringsAsFactors = FALSE)
wind <- read.csv("dfwinddatetime_VOC.csv")
```
# Benzene Analysis
## Convert timestamps into datetime format in the benzene and wind datasets
```{r}
#Convert vector into characters
benzene$Datetime <- as.character(benzene$startTimestamp)
wind$Datetime <- as.character(wind$Datetime)
```
```{r}
#Convert vector of characters into POSIXct
benzene$Datetime <- as.POSIXct(benzene$startTimestamp, format="%Y-%m-%d%H:%M")
benzene$Datetime[1:6]

wind$Datetime <- as.POSIXct(wind$Datetime, format="%Y-%m-%d%H:%M")
wind$Datetime[1:6]
```
```{r}
#get rid of timezone 
strip.tz <- function(dt) {
  fmt <- "%Y-%m-%d %H:%M:%S"
  strptime(strftime(dt, format = fmt, tz=""), format = fmt, tz="UTC")
}

benzene$Datetime <- strip.tz(benzene$Datetime)
wind$Datetime <- strip.tz(wind$Datetime)
```
```{r}
#Look to see if it worked
str(benzene$Datetime)
str(wind$Datetime)
```
### Round datetime in VOC data to hours
*We are doing this in order to match the benzene data with the wind data*
```{r}
#Round the benzene datetime to hours
benzene$Datetime <- round(benzene$Datetime,"hours")
```
```{r}
#See if it worked
benzene$Datetime[1:6]
wind$Datetime[1:6]
```
## Merge the Benzene and Wind Data
```{r}
#perform the merge by datetime
omni_wind <- merge(wind, benzene, by = 'Datetime')
#see if it worked
head(omni_wind)
```
```{r}
#make sure Datetime stays in POSIXct format
omni_wind$Datetime <- as.POSIXct(omni_wind$Datetime, format="%Y-%m-%d%H:%M")
```
### Add tube average to the data frame for reference
```{r}
#add in tube average to make a reference line
omni_wind$tube_average_round2 <- NA
omni_wind[which(omni_wind$Datetime >="2020-04-05 15:39:34" & omni_wind$Datetime<="2020-04-12 16:35:45"),c("tube_average_round2")] <- 0.203

omni_wind$tube_average_round3 <- NA
omni_wind[which(omni_wind$Datetime>="2020-04-25 11:01:42" & omni_wind$Datetime<="2020-05-02 05:41:04"),c("tube_average_round3")] <- 0.196

#Note: this can also just be added as a line in the graph
```
## Get rid of NAs for Windspeed
```{r}
omni_wind <- omni_wind %>% drop_na(mean, ws)
```
## Delete Duplicate Rows
```{r}
#delete duplicate rows
omni_wind2 <- unique( omni_wind )
```
## Replace NAs for Benzene with Half of the MDL 
```{r}
#make sure concentration is numeric
omni_wind2$benzene_concentration_ppb <- as.numeric(omni_wind2$benzene_concentration_ppb)
#turn "N/A" into 0.7 ppb
omni_wind2[is.na(omni_wind2$benzene_concentration_ppb),c('benzene_concentration_ppb')] <- 0.7
```
## Find the Benzene Average
```{r}
#what is the omniscent average for benzene?
mean(omni_wind2$benzene_concentration_ppb)
```
# Benzene Visualizations
## Plot the Benzene and Wind Direction Data
```{r}
omni_wind2$Wind <- factor(omni_wind2$Wind, levels = c("North", "Northeast", "East", "Southeast", "South", "Southwest", "West", "Northwest"))

p<- ggplot(omni_wind2, aes(x = Wind, y = benzene_concentration_ppb, color=Wind)) + geom_point(aes(size=2)) +theme_classic() + ggtitle("Omniscent Benzene Concentrations over Time and Wind Direction") + xlab("Wind Direction") + ylab("Benzene Concentration (ppb)") + geom_hline(yintercept=mean(omni_wind2$benzene_concentration_ppb), linetype="solid", color = "black")+ theme(text=element_text(family="sans", size=14)) + guides(size = FALSE)+ scale_x_discrete(guide = guide_axis(angle = 45), limits = c("North", "Northeast", "East", "Southeast", "South", "Southwest", "West", "Northwest"))
p<- update_labels(p, list(colour="Benzene Concentration (ppb)"))
p
```
## Plot the Benzene Data Colored According to Wind Direction
```{r}
p1 <- ggplot(omni_wind2, aes(Datetime, color=Measurements)) + geom_point(aes(y = benzene_concentration_ppb, color = "Omniscent" )) + geom_line(aes(y=tube_average_round2, color = "Tube Average", size=0.5)) + geom_line(aes(y=tube_average_round3, color = "Tube Average", size=0.5)) + geom_hline(yintercept= mean(omni_wind2$benzene_concentration_ppb), linetype="solid", color = "black") + theme_classic() + ggtitle("Omniscent and ATD Tube Benzene Concentrations over Time") + xlab("Date") + ylab("Benzene Concentration (ppb)") + guides(size = FALSE) 

p1+ labs(fill ="Legend")+ theme(text=element_text(family="sans", size=14))
```
## Benzene Wind Rose
### Create Wind Rose for All Benzene Measurements
```{r}
omni_wind2$wd <- omni_wind2$mean

windRose(omni_wind2, ws = "ws", wd = "mean", ws2 = NA, wd2 = NA, 
  ws.int = 2, angle = 30, type = "default", bias.corr = TRUE, cols
  = "default", grid.line = NULL, width = 1, seg = NULL, auto.text 
  = TRUE, breaks = 4, offset = 10, normalise = FALSE, max.freq = 
  NULL, paddle = TRUE, key.header = NULL, key.footer = "mph", 
  key.position = "bottom", key = TRUE, dig.lab = 5, statistic = 
  "prop.count", pollutant = 'benzene_concentration_ppb', annotate = TRUE, angle.scale = 
  315, border = NA)

pollutionRose(omni_wind2, pollutant = "benzene_concentration_ppb", key.footer = "Benzene Concentration (ppb)")
```
### Create Wind Rose for All Benzene Measurements over 1.4 ppb
```{r}

omni_wind3 <- omni_wind2[which((omni_wind2$benzene_concentration_ppb) > 1.4),]

windRose(omni_wind3, ws = "ws", wd = "mean", ws2 = NA, wd2 = NA, 
  ws.int = 2, angle = 30, type = "default", bias.corr = TRUE, cols
  = "default", grid.line = NULL, width = 1, seg = NULL, auto.text 
  = TRUE, breaks = 3, offset = 10, normalise = FALSE, max.freq = 
  NULL, paddle = TRUE, key.header = NULL, key.footer = "mph", 
  key.position = "bottom", key = TRUE, dig.lab = 5, statistic = 
  "prop.count", pollutant = 'benzene_concentration_ppb', annotate = TRUE, angle.scale = 
  315, border = NA)
```
### Create Pollution Rose for All Benzene Measurements over 1.4 ppb
```{r}
pollutionRose(omni_wind3, pollutant = "benzene_concentration_ppb", key.footer = "Benzene Concentration (ppb)", main = "Benzene Measurements over 1.4 ppb and Wind Direction")
```

# Benzene Analysis
## Convert timestamps into datetime format in the benzene and wind datasets
```{r}
toluene$Datetime <- as.character(toluene$startTimestamp)
wind$Datetime <- as.character(wind$Datetime)
```
```{r}
toluene$Datetime <- as.POSIXct(toluene$startTimestamp, format="%Y-%m-%d%H:%M")
toluene$Datetime[1:6]

wind$Datetime <- as.POSIXct(wind$Datetime, format="%Y-%m-%d%H:%M")
wind$Datetime[1:6]
```
```{r}
#get rid of timezone 
strip.tz <- function(dt) {
  fmt <- "%Y-%m-%d %H:%M:%S"
  strptime(strftime(dt, format = fmt, tz=""), format = fmt, tz="UTC")
}

toluene$Datetime <- strip.tz(toluene$Datetime)
wind$Datetime <- strip.tz(wind$Datetime)
```
```{r}
str(toluene$Datetime)
str(wind$Datetime)
```
### Round datetime in VOC data to hours
*We are doing this in order to match the benzene data with the wind data*
```{r}
toluene$Datetime <- round(toluene$Datetime,"hours")
```
```{r}
toluene$Datetime[1:6]
wind$Datetime[1:6]
```
## Merge the Benzene and Wind Data 
```{r}
omni_windt <- merge(wind, toluene, by = 'Datetime')
head(omni_windt)
```
```{r}
#make sure Datetime stays in POSIXct format
omni_windt$Datetime <- as.POSIXct(omni_windt$Datetime, format="%Y-%m-%d%H:%M")
```
```{r}
#add in tube averages
omni_windt$tube_average_round2 <- NA
omni_windt[which(omni_windt$Datetime >="2020-04-05 15:39:34" & omni_windt$Datetime<="2020-04-12 16:35:45"),c("tube_average_round2")] <- 0.203

omni_windt$tube_average_round3 <- NA
omni_windt[which(omni_windt$Datetime>="2020-04-25 11:01:42" & omni_windt$Datetime<="2020-05-02 05:41:04"),c("tube_average_round3")] <- 0.361
```
```{r}
omni_windt <- omni_windt %>% drop_na(mean, ws)
```


```{r}
#delete duplicate rows
omni_windt2 <- unique( omni_windt )
#make sure toluene concentration is numeric
omni_windt2$toluene_Concentration..ppb. <- as.numeric(omni_windt2$toluene_Concentration..ppb.)
#turn "N/A" into 1/2 MDL
omni_windt2[is.na(omni_windt2$toluene_Concentration..ppb.),c('toluene_Concentration..ppb.')] <- 0.2

```
```{r}
#what is the omniscent average for toluene?
mean(omni_windt2$toluene_Concentration..ppb.)
``` 

```{r}
ggplot(omni_windt2, aes(x = Datetime, y = toluene_Concentration..ppb., color=Wind)) + geom_point(aes(size=2)) +theme_classic() + ggtitle("Omniscent Toluene Concentrations over Time and Wind Direction") + xlab("Date") + ylab("Toluene Concentration (ppb)") + geom_hline(yintercept=mean(omni_windt2$toluene_Concentration..ppb..with.MDL), linetype="solid", color = "black")+ theme(text=element_text(family="sans", size=14)) + guides(size = FALSE)+ scale_color_manual(values = c("East" = "#450acb","North"="#2e0787", "Northeast"="#3908a9", "Northwest"="steelblue", "South"="#6729f3", "Southeast"="#500ced", "Southwest"="#7f4bf5", "West"="#976df7"))
```
```{r}
omni_windt2$Wind <- factor(omni_windt2$Wind, levels = c("North", "Northeast", "East", "Southeast", "South", "Southwest", "West", "Northwest"))

legend_title <- "OMG My Title"

p<- ggplot(omni_windt2, aes(x = Wind, y = toluene_Concentration..ppb., color=Wind)) + geom_point(aes(size=2)) +theme_classic() + ggtitle("Omniscent Toluene Concentrations over Time and Wind Direction") + xlab("Wind Direction") + ylab("Toluene Concentration (ppb)") + geom_hline(yintercept=mean(omni_windt2$toluene_Concentration..ppb..with.MDL), linetype="solid", color = "black")+ theme(text=element_text(family="sans", size=14)) + guides(size = FALSE)+ scale_x_discrete(guide = guide_axis(angle = 45), limits = c("North", "Northeast", "East", "Southeast", "South", "Southwest", "West", "Northwest"))
p<- update_labels(p, list(colour="Toluene Concentration (ppb)"))
p
```

```{r}
ggplot(omni_windt2, aes(Datetime, color=Measurements)) + geom_point(aes(y = toluene_Concentration..ppb., color = "Omniscent")) + geom_line(aes(y=tube_average_round2, color = "Tube Average", size=0.5)) + geom_line(aes(y=tube_average_round3, color = "Tube Average", size=0.5)) + theme_classic() + ggtitle("Omniscent and ATD Tube Benzene Concentrations over Time") + xlab("Date") + ylab("Toluene Concentration (ppb)") + guides(size = FALSE)+ geom_hline(yintercept=mean(omni_windt2$toluene_Concentration..ppb..with.MDL), linetype="solid", color = "black")+ theme(text=element_text(family="sans", size=14))
```

```{r}
omni_windt2$wd <- omni_windt2$mean
windRose(omni_windt2, ws = "ws", wd = "mean", ws2 = NA, wd2 = NA, 
  ws.int = 2, angle = 30, type = "default", bias.corr = TRUE, cols
  = "default", grid.line = NULL, width = 1, seg = NULL, auto.text 
  = TRUE, breaks = 4, offset = 10, normalise = FALSE, max.freq = 
  NULL, paddle = TRUE, key.header = NULL, key.footer = "mph", 
  key.position = "bottom", key = TRUE, dig.lab = 5, statistic = 
  "prop.count", pollutant = 'toluene_Concentration..ppb.', annotate = TRUE, angle.scale = 
  315, border = NA)

pollutionRose(omni_windt2, pollutant = "toluene_Concentration..ppb.", key.footer = "Toluene Concentration (ppb)")
```

```{r}
omni_windt3 <- omni_windt2[which((omni_windt2$toluene_Concentration..ppb..with.MDL) > 0.4),]

windRose(omni_windt3, ws = "ws", wd = "mean", ws2 = NA, wd2 = NA, 
  ws.int = 2, angle = 30, type = "default", bias.corr = TRUE, cols
  = "default", grid.line = NULL, width = 1, seg = NULL, auto.text 
  = TRUE, breaks = 4, offset = 10, normalise = FALSE, max.freq = 
  NULL, paddle = TRUE, key.header = NULL, key.footer = "mph", 
  key.position = "bottom", key = TRUE, dig.lab = 5, statistic = 
  "prop.count", pollutant = 'toluene_Concentration..ppb.', annotate = TRUE, angle.scale = 
  315, border = NA)

pollutionRose(omni_windt3, pollutant = "toluene_Concentration..ppb.", key.footer = "Toluene Concentration (ppb)", main = "Positive Toluene Concentrations and Wind Direction")
```



